---
title: "Champion Stats Analysis"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Install necessary packages (if not already installed)
# install.packages(c("rvest", "tidyverse", "stringr", "purrr"))
# install.packages("GGally")
# install.packages("plotly")
# install.packages("kableExtra")
# install.packages("randomForest")
# install.packages("DT")

library(plotly)
library(GGally)
library(rvest)
library(tidyverse)
library(dplyr)
library(purrr)
library(stringr)
library(jsonlite)
library(kableExtra)
library(tidyr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(jsonlite)
library(tibble)
library(mgcv)
library(randomForest)
library(DT)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Wayback Machine Archive Data
#############################################

# Define the TFTTactics.gg Champions URL for Set 12
url_tfttactics_set12 <- "https://web.archive.org/web/20240927225144/https://tftactics.gg/champions/"
page_tfttactics_set12 <- read_html(url_tfttactics_set12)

# Extract champion names for Set 12
champion_names_set12 <- page_tfttactics_set12 %>%
  html_nodes(".character-name") %>% 
  html_text(trim = TRUE)

# Extract trait names (global list) for Set 12
trait_names_set12 <- page_tfttactics_set12 %>%
  html_nodes(".dropdown-item img") %>%
  html_attr("alt") 

# Initialize the tibbles with proper column names for Set 12
champion_info_set12 <- tibble(
  champion = character(),
  stats = list(),
  abilities = list(),
  traits = list(),
  trait_breakpoints = list()
)

trait_info_set12 <- tibble(
  champion = character(),
  trait_name = character(),
  trait_type = character(),
  trait_description = character()
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Wayback Machine Archive Data
#############################################

# Define the TFTTactics.gg Champions URL for Set 13
url_tfttactics_set13 <- "https://web.archive.org/web/20250114220710/https://tftactics.gg/champions/"
page_tfttactics_set13 <- read_html(url_tfttactics_set13)

# Extract champion names for Set 13
champion_names_set13 <- page_tfttactics_set13 %>%
  html_nodes(".character-name") %>% 
  html_text(trim = TRUE)

# Extract trait names (global list) for Set 13
trait_names_set13 <- page_tfttactics_set13 %>%
  html_nodes(".dropdown-item img") %>%
  html_attr("alt") 

# Initialize the tibbles with proper column names for Set 13
champion_info_set13 <- tibble(
  champion = character(),
  stats = list(),
  abilities = list(),
  traits = list(),
  trait_breakpoints = list()
)

trait_info_set13 <- tibble(
  champion = character(),
  trait_name = character(),
  trait_type = character(),
  trait_description = character()
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Common Functions (Used for both sets)
#############################################

# Function to safely read the page with error handling
safe_read_html <- function(url) {
  tryCatch(
    {
      read_html(url)
    },
    error = function(e) {
      message("Error reading URL: ", url)
      message("Error: ", e$message)
      NULL
    }
  )
}

# Function to extract champion data using regex patterns
extract_champion_data <- function(page, trait_names) {
  # Get raw HTML text
  raw_text <- page %>% html_text()
  
  # Extract champion name from the title
  champion_name <- page %>% 
    html_node("title") %>% 
    html_text() %>% 
    str_extract("^[^·]+") %>% 
    str_trim()
  
  # stats EXTRACTION
  stats <- tibble(stat = character(), value = character())
  stat_patterns <- list(
    "Cost" = "Cost:\\s*(\\d+)",
    "Health" = "Health:\\s*([\\d\\s\\/]+)",
    "Mana" = "Mana:\\s*(\\d+)",
    "Armor" = "Armor:\\s*(\\d+)",
    "MR" = "MR:\\s*(\\d+)",
    "DPS" = "DPS:\\s*([\\d\\s\\/]+)",
    "Damage" = "Damage:\\s*([\\d\\s\\/]+)",
    "Atk Spd" = "Atk Spd:\\s*([\\d\\.]+)",
    "Crit Rate" = "Crit Rate:\\s*([\\d\\.%]+)",
    "Range" = "Range:\\s*(\\d+)"
  )
  
  for (stat_name in names(stat_patterns)) {
    value <- str_extract(raw_text, stat_patterns[[stat_name]])
    if (!is.na(value)) {
      clean_value <- str_replace(value, paste0(stat_name, ":\\s*"), "") %>% str_trim()
      stats <- stats %>% add_row(stat = stat_name, value = clean_value)
    }
  }
  
  # ABILITIES EXTRACTION
  ability_section <- str_extract(raw_text, "Abilities(.+?)(?=[A-Z][a-z]+(?:Origin|Class)|Synergies|Privacy Policy)")
  ability_name <- str_extract(ability_section, "(?<=Abilities)\\s*([A-Za-z\\s]+?)(?=Active|Passive)") %>% str_trim()
  ability_type <- str_extract(ability_section, "(Active|Passive)")
  ability_mana <- str_extract(ability_section, "\\d+\\s*\\/\\s*\\d+")

  abilities <- tibble(
    name = ability_name,
    type = ability_type,
    mana = ability_mana,
    description = NA_character_
  )

  # TRAITS EXTRACTION
  champion_traits <- character()
  for (trait in trait_names) {
    if (str_detect(raw_text, fixed(trait))) {
      champion_traits <- c(champion_traits, trait)
    }
  }

  traits <- tibble(name = character(), type = character(), description = character())
  trait_breakpoints <- tibble(trait_name = character(), breakpoint_value = character(), breakpoint_description = character())

  for (trait_name in champion_traits) {
    next_traits <- setdiff(trait_names, trait_name)
    next_part <- if (length(next_traits) > 0) paste(next_traits, collapse = "|") else ""

    end_pattern <- if (nchar(next_part) > 0) {
      paste0("(?=", next_part, "|Synergies)")
    } else {
      "(?=Synergies)"
    }

    trait_pattern <- paste0(trait_name, "(.+?)", end_pattern)
    trait_section <- str_extract(raw_text, trait_pattern)

    if (!is.na(trait_section)) {
      trait_type <- str_extract(trait_section, "(Origin|Class)")

      desc_pattern <- paste0("(?<=", trait_type, ")\\s*(.+?)(?=(", 
                             if(nchar(next_part) > 0) paste0(next_part, "|") else "", 
                             "Synergies|\\d+%|\\d+\\s*AD))")

      trait_desc <- str_extract(trait_section, desc_pattern) %>% str_trim()

      traits <- traits %>% add_row(
        name = trait_name,
        type = trait_type,
        description = trait_desc
      )

      breakpoint_pattern <- "(\\d+)(.+?)(?=\\d+|$)"
      breakpoint_matches <- str_extract_all(trait_section, breakpoint_pattern)[[1]]
      for (match in breakpoint_matches) {
        bp_value <- str_extract(match, "^\\d+") %>% str_trim()
        bp_desc <- str_replace(match, "^\\d+", "") %>% str_trim()
        trait_breakpoints <- trait_breakpoints %>% add_row(
          trait_name = trait_name,
          breakpoint_value = bp_value,
          breakpoint_description = bp_desc
        )
      }
    }
  }

  return(list(
    champion = champion_name,
    stats = stats,
    abilities = abilities,
    traits = traits,
    trait_breakpoints = trait_breakpoints
  ))
}

# Function to extract champion stats for analysis
extract_champion_stats <- function(champion_name, stats_tibble) {
  # Extract Cost, Health, DPS, and Damage stats
  cost <- stats_tibble %>% filter(stat == "Cost") %>% pull(value) %>% as.numeric()
  health <- stats_tibble %>% filter(stat == "Health") %>% pull(value)
  dps <- stats_tibble %>% filter(stat == "DPS") %>% pull(value)
  damage <- stats_tibble %>% filter(stat == "Damage") %>% pull(value)
  
  # Split the health, dps, and damage values by " / " to separate star levels
  health_values <- as.numeric(strsplit(health, " / ")[[1]])
  dps_values <- as.numeric(strsplit(dps, " / ")[[1]])
  damage_values <- as.numeric(strsplit(damage, " / ")[[1]])

  # Ensure that the number of values for health, dps, and damage match across star levels
  if (length(health_values) != length(dps_values) | length(dps_values) != length(damage_values)) {
    stop("Mismatch in the number of star levels for health, dps, and damage.")
  }
  
  # Create a tibble with the expanded stats
  stats_df <- tibble(
    champion = rep(champion_name, length(health_values)),
    star_level = 1:length(health_values),
    cost = rep(cost, length(health_values)),
    health = health_values,
    dps = dps_values,
    damage = damage_values
  )
  
  return(stats_df)
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Extract All Champion Data
#############################################

extract_all_champion_data_set12 <- function(champion_names) {
  # Reset global variables
  champion_info_set12 <<- tibble(
    champion = character(),
    stats = list(),
    abilities = list(),
    traits = list(),
    trait_breakpoints = list()
  )

  trait_info_set12 <<- tibble(
    champion = character(),
    trait_name = character(),
    trait_type = character(),
    trait_description = character()
  )

  base_url <- "https://web.archive.org/web/20240927225144/https://tftactics.gg/champions/"

  for (champion_name in champion_names) {
    champion_url <- paste0(base_url, gsub(" ", "_", champion_name), "/")
    champion_page <- safe_read_html(champion_url)

    if (is.null(champion_page)) {
      next
    }

    champion_data <- extract_champion_data(champion_page, trait_names_set12)

    champion_info_set12 <<- champion_info_set12 %>% add_row(
      champion = champion_data$champion,
      stats = list(champion_data$stats),
      abilities = list(champion_data$abilities),
      traits = list(champion_data$traits),
      trait_breakpoints = list(champion_data$trait_breakpoints)
    )

    for (trait in champion_data$traits$name) {
      trait_info_set12 <<- trait_info_set12 %>% add_row(
        champion = champion_data$champion,
        trait_name = trait,
        trait_type = champion_data$traits$type[champion_data$traits$name == trait],
        trait_description = champion_data$traits$description[champion_data$traits$name == trait]
      )
    }
  }

  return(list(champion_info = champion_info_set12, trait_info = trait_info_set12))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Extract All Champion Data
#############################################

extract_all_champion_data_set13 <- function(champion_names) {
  # Reset global variables
  champion_info_set13 <<- tibble(
    champion = character(),
    stats = list(),
    abilities = list(),
    traits = list(),
    trait_breakpoints = list()
  )

  trait_info_set13 <<- tibble(
    champion = character(),
    trait_name = character(),
    trait_type = character(),
    trait_description = character()
  )

  base_url <- "https://tftactics.gg/champions/"

  for (champion_name in champion_names) {
    champion_url <- paste0(base_url, gsub(" ", "_", champion_name), "/")
    champion_page <- safe_read_html(champion_url)

    if (is.null(champion_page)) {
      next
    }

    champion_data <- extract_champion_data(champion_page, trait_names_set13)

    champion_info_set13 <<- champion_info_set13 %>% add_row(
      champion = champion_data$champion,
      stats = list(champion_data$stats),
      abilities = list(champion_data$abilities),
      traits = list(champion_data$traits),
      trait_breakpoints = list(champion_data$trait_breakpoints)
    )

    for (trait in champion_data$traits$name) {
      trait_info_set13 <<- trait_info_set13 %>% add_row(
        champion = champion_data$champion,
        trait_name = trait,
        trait_type = champion_data$traits$type[champion_data$traits$name == trait],
        trait_description = champion_data$traits$description[champion_data$traits$name == trait]
      )
    }
  }

  return(list(champion_info = champion_info_set13, trait_info = trait_info_set13))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Extract Data for Both Sets
#############################################
# Extract data for both sets
# all_data_set12 <- extract_all_champion_data_set12(champion_names_set12)
# all_data_set13 <- extract_all_champion_data_set13(champion_names_set13)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read the JSON files
champion_info_set12 <- fromJSON("data/champion_info_set12.json")
trait_info_set12 <- fromJSON("data/trait_info_set12.json")

# Initialize all_data_set_12 as an empty data frame
all_data_set12 <- list()

# Add the JSON data as columns to the new data frame
all_data_set12$champion_info <- champion_info_set12
all_data_set12$trait_info <- trait_info_set12
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read the JSON files
champion_info_set13 <- fromJSON("data/champion_info_set13.json")
trait_info_set13 <- fromJSON("data/trait_info_set13.json")

# Initialize all_data_set_13 as an empty data frame
all_data_set13 <- list()

# Add the JSON data as columns to the new data frame
all_data_set13$champion_info <- champion_info_set13
all_data_set13$trait_info <- trait_info_set13
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Process Data for Both Sets
#############################################
# Clean champion names for Set 12
all_data_set12$trait_info <- all_data_set12$trait_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide")) %>%
  filter(!(trait_name == "Portal" & is.na(trait_type)))

all_data_set12$champion_info <- all_data_set12$champion_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

# Clean champion names for Set 13
all_data_set13$trait_info <- all_data_set13$trait_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

all_data_set13$champion_info <- all_data_set13$champion_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

# print(all_data_set12$trait_info)
# print(all_data_set12$champion_info)
# print(all_data_set13$trait_info)
# print(all_data_set13$champion_info)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Extract All Winrate Data
#############################################

# Read JSON
url_set12 <- "https://web.archive.org/web/20240917151936/https://tftactics-data.kda.gg/meta/report"
report_data_set12 <- fromJSON(url_set12)

# Go to the summary-section
summary_section_set12 <- report_data_set12$data$`summary-section`

# Helper function to clean a single name
clean_champion_name_set12 <- function(x) {
  x <- gsub("^[0-9]-", "", x)
  x <- gsub("^tft12_", "", x)
  x <- gsub("-[0-9]+$", "", x)
  x <- tolower(x)
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x
}

# Build the dataframe
comp_df_set12 <- tibble(
  win_rate = summary_section_set12$win_rate,
  top4_rate = summary_section_set12$top4_rate,
  champions = sapply(summary_section_set12$composition, function(champs_vec) {
    cleaned_champions <- sapply(champs_vec, clean_champion_name_set12)
    paste(cleaned_champions, collapse = ", ")
  })
)

# Separate Champions from Traits
comp_df_clean_set12 <- comp_df_set12 %>%
  mutate(champions = str_split(champions, ",\\s*")) %>%
  rowwise() %>%
  mutate(
    champions_list = list(champions[champions %in% champion_names_set12]),
    traits_list = list(champions[!(champions %in% champion_names_set12)])
  ) %>%
  ungroup() %>%
  mutate(
    champions = sapply(champions_list, function(x) paste(x, collapse = ", ")),
    traits = sapply(traits_list, function(x) paste(x, collapse = ", "))
  ) %>%
  select(win_rate, top4_rate, champions, traits)

# print(comp_df_clean_set12)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Extract All Winrate Data
#############################################

# Download compressed file
temp_file_set13 <- tempfile(fileext = ".gz")
download.file(
  url = "https://web.archive.org/web/20250113142238/https://tftactics-data.kda.gg/meta/report",
  destfile = temp_file_set13,
  mode = "wb"
)

# Decompress using gzcon
con_set13 <- gzcon(file(temp_file_set13, "rb"))
text_set13 <- readLines(con_set13, warn = FALSE)
close(con_set13)

# Parse the text as JSON
report_data_set13 <- fromJSON(paste(text_set13, collapse = ""))

# Go to the summary-section
summary_section_set13 <- report_data_set13$data$`summary-section`

# Helper function to clean a single name
clean_champion_name_set13 <- function(x) {
  x <- gsub("^[0-9]-", "", x)
  x <- gsub("^tft13_", "", x)
  x <- gsub("-[0-9]+$", "", x)
  x <- tolower(x)
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x
}

# Build the dataframe
comp_df_set13 <- tibble(
  win_rate = summary_section_set13$win_rate,
  top4_rate = summary_section_set13$top4_rate,
  champions = sapply(summary_section_set13$composition, function(champs_vec) {
    cleaned_champions <- sapply(champs_vec, clean_champion_name_set13)
    paste(cleaned_champions, collapse = ", ")
  })
)

# Separate Champions from Traits
comp_df_clean_set13 <- comp_df_set13 %>%
  mutate(champions = str_split(champions, ",\\s*")) %>%
  rowwise() %>%
  mutate(
    champions_list = list(champions[champions %in% champion_names_set13]),
    traits_list = list(champions[!(champions %in% champion_names_set13)])
  ) %>%
  ungroup() %>%
  mutate(
    champions = sapply(champions_list, function(x) paste(x, collapse = ", ")),
    traits = sapply(traits_list, function(x) paste(x, collapse = ", "))
  ) %>%
  select(win_rate, top4_rate, champions, traits)

# print(comp_df_clean_set13)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 12 and Set 13 Champion and Trait Stats
#############################################

# Explode champions and traits into separate rows

# For Set 12 - explode champions
comp_df_clean_set12_exploded <- comp_df_clean_set12 %>%
  mutate(
    champions = str_split(champions, ",\\s*")
  ) %>%
  unnest(champions) %>%
  mutate(
    champions = str_trim(champions)
  )

# For Set 13 - explode champions
comp_df_clean_set13_exploded <- comp_df_clean_set13 %>%
  mutate(
    champions = str_split(champions, ",\\s*")
  ) %>%
  unnest(champions) %>%
  mutate(
    champions = str_trim(champions)
  )

# For Set 12 - explode traits
comp_df_clean_set12_traits_exploded <- comp_df_clean_set12 %>%
  mutate(
    traits = str_split(traits, ",\\s*")
  ) %>%
  unnest(traits) %>%
  mutate(
    traits = str_trim(traits)
  )

# For Set 13 - explode traits
comp_df_clean_set13_traits_exploded <- comp_df_clean_set13 %>%
  mutate(
    traits = str_split(traits, ",\\s*")
  ) %>%
  unnest(traits) %>%
  mutate(
    traits = str_trim(traits)
  )

# Summarize Champion Stats
champion_summary_set12 <- comp_df_clean_set12_exploded %>%
  group_by(champions) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

champion_summary_set13 <- comp_df_clean_set13_exploded %>%
  group_by(champions) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# Summarize Trait Stats
trait_summary_set12 <- comp_df_clean_set12_traits_exploded %>%
  group_by(traits) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

trait_summary_set13 <- comp_df_clean_set13_traits_exploded %>%
  group_by(traits) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# champion_summary_set12
# champion_summary_set13
# trait_summary_set12
# trait_summary_set13
```

# Champion Stat Distribution (No Win Rate)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 12 Champion stats
#############################################

# Prepare champion stats for analysis (Set 12)
all_champion_stats_set12 <- bind_rows(
  lapply(1:nrow(all_data_set12$champion_info), function(i) {
    champion_name <- all_data_set12$champion_info$champion[i]
    champion_stats <- all_data_set12$champion_info$stats[[i]]
    
    # Extract and combine stats
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
      # return(NULL)
    })
  })
)

# Remove any NULL entries
all_champion_stats_set12 <- all_champion_stats_set12 %>% filter(!is.na(champion))

# Arrange champions by various attributes (Set 12)
arranged_champions_set12 <- all_champion_stats_set12 %>%
  arrange(champion, star_level, cost)

# Compute averages for health, dps, and damage grouped by cost and star level (Set 12)
averaged_champions_set12 <- arranged_champions_set12 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping after summarisation
  )

# Sort tables by different metrics (Set 12)
ordered_by_health_set12 <- averaged_champions_set12 %>%
  arrange(avg_health) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_dps_set12 <- averaged_champions_set12 %>%
  arrange(avg_dps) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_damage_set12 <- averaged_champions_set12 %>%
  arrange(avg_damage) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

# Create kable tables for Set 12
kable_health_set12 <- kable(ordered_by_health_set12, 
                          caption = "Set 12: Champions Ordered by Average Health",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

kable_dps_set12 <- kable(ordered_by_dps_set12, 
                       caption = "Set 12: Champions Ordered by Average DPS",
                       col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                       format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

kable_damage_set12 <- kable(ordered_by_damage_set12, 
                          caption = "Set 12: Champions Ordered by Average Damage",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

# print(kable_health_set12)
# print(kable_dps_set12)
# print(kable_damage_set12)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 13 Champion stats
#############################################

# Prepare champion stats for analysis (Set 13)
all_champion_stats_set13 <- bind_rows(
  lapply(1:nrow(all_data_set13$champion_info), function(i) {
    champion_name <- all_data_set13$champion_info$champion[i]
    champion_stats <- all_data_set13$champion_info$stats[[i]]
    
    # Extract and combine stats
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
      # return(NULL)
    })
  })
)

# Remove any NULL entries
all_champion_stats_set13 <- all_champion_stats_set13 %>% filter(!is.na(champion))

# Arrange champions by various attributes (Set 13)
arranged_champions_set13 <- all_champion_stats_set13 %>%
  arrange(champion, star_level, cost)

# Compute averages for health, dps, and damage grouped by cost and star level (Set 13)
averaged_champions_set13 <- arranged_champions_set13 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping after summarisation
  )

# Sort tables by different metrics (Set 13)
ordered_by_health_set13 <- averaged_champions_set13 %>%
  arrange(avg_health) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_dps_set13 <- averaged_champions_set13 %>%
  arrange(avg_dps) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_damage_set13 <- averaged_champions_set13 %>%
  arrange(avg_damage) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

# Create kable tables for Set 13
kable_health_set13 <- kable(ordered_by_health_set13, 
                          caption = "Set 13: Champions Ordered by Average Health",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

kable_dps_set13 <- kable(ordered_by_dps_set13, 
                       caption = "Set 13: Champions Ordered by Average DPS",
                       col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                       format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

kable_damage_set13 <- kable(ordered_by_damage_set13, 
                          caption = "Set 13: Champions Ordered by Average Damage",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

# print(kable_health_set13)
# print(kable_dps_set13)
# print(kable_damage_set13)
```

## Set 12: Stats by Cost and Star Level

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8}
#############################################
# Create Interactive Visualizations for Set 12
#############################################

# Scatterplots for Set 12
set12_health_scatter <- ggplot(arranged_champions_set12, aes(x = as.factor(cost), y = health, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "Health by Cost and Star Level", x = "Cost", y = "Health", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

set12_damage_scatter <- ggplot(arranged_champions_set12, aes(x = as.factor(cost), y = damage, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "Damage by Cost and Star Level", x = "Cost", y = "Damage", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

set12_dps_scatter <- ggplot(arranged_champions_set12, aes(x = as.factor(cost), y = dps, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "DPS by Cost and Star Level", x = "Cost", y = "DPS", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

# Display interactive plots separately
plotly::ggplotly(set12_health_scatter)
plotly::ggplotly(set12_damage_scatter)
plotly::ggplotly(set12_dps_scatter)
```

There is a clear trend across all star levels: as the cost increases, all three stats (Health, Damage, and DPS) increase noticeably.

## Set 13: Stats by Cost and Star Level

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8}
#############################################
# Create Interactive Visualizations for Set 13
#############################################

# Scatterplots for Set 13
set13_health_scatter <- ggplot(arranged_champions_set13, aes(x = as.factor(cost), y = health, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "Health by Cost and Star Level", x = "Cost", y = "Health", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

set13_damage_scatter <- ggplot(arranged_champions_set13, aes(x = as.factor(cost), y = damage, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "Damage by Cost and Star Level", x = "Cost", y = "Damage", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

set13_dps_scatter <- ggplot(arranged_champions_set13, aes(x = as.factor(cost), y = dps, color = as.factor(star_level))) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", aes(group = as.factor(star_level)), se = FALSE) +
  labs(title = "DPS by Cost and Star Level", x = "Cost", y = "DPS", color = "Star Level") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")

# Display interactive plots separately
plotly::ggplotly(set13_health_scatter)
plotly::ggplotly(set13_damage_scatter)
plotly::ggplotly(set13_dps_scatter)
```

The trend for Set 13 is similar to that of Set 12, but the gradient is steeper and the maximum values are much higher. This is expected because Set 13 includes 6-cost champions, which are very rare and have exceptionally high stats to compensate for their high cost and rarity. An outlier is observed in the Damage and DPS graphs, where a 3-cost, 2-star unit has very low damage and DPS—possibly explained by its role as a tank with high health and low damage.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Side-by-Side Comparison - Creating Combined Dataset
#############################################

# Extract champion stats for Set 12
all_champion_stats_set12 <- bind_rows(
  lapply(1:nrow(all_data_set12$champion_info), function(i) {
    champion_name <- all_data_set12$champion_info$champion[i]
    champion_stats <- all_data_set12$champion_info$stats[[i]]
    
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
    })
  })
) %>% filter(!is.na(champion))

# Extract champion stats for Set 13
all_champion_stats_set13 <- bind_rows(
  lapply(1:nrow(all_data_set13$champion_info), function(i) {
    champion_name <- all_data_set13$champion_info$champion[i]
    champion_stats <- all_data_set13$champion_info$stats[[i]]
    
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
    })
  })
) %>% filter(!is.na(champion))

# Arrange champions for both sets
arranged_champions_set12 <- all_champion_stats_set12 %>%
  arrange(champion, star_level, cost)

arranged_champions_set13 <- all_champion_stats_set13 %>%
  arrange(champion, star_level, cost)

# Calculate average stats by cost and star level for both sets
averaged_champions_set12 <- arranged_champions_set12 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'
  )

averaged_champions_set13 <- arranged_champions_set13 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'
  )

# Create a combined dataset for comparison without merging based on keys
set12_avg_summary <- averaged_champions_set12 %>%
  mutate(set = "Set 12")

set13_avg_summary <- averaged_champions_set13 %>%
  mutate(set = "Set 13")

# Bind the rows without merging or matching keys
combined_sets <- bind_rows(set12_avg_summary, set13_avg_summary)
```

## Set 12 vs Set 13: Average stats Comparison by Cost and Star Level

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8}
#############################################
# Interactive Bar Chart Comparisons
#############################################

# Create comparison visualizations for health, DPS, and damage
health_comparison <- ggplot(combined_sets, aes(x = as.factor(cost), y = avg_health, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Average Health Comparison", x = "Cost", y = "Average Health") +
  theme_minimal() +
  theme(legend.title = element_blank())

dps_comparison <- ggplot(combined_sets, aes(x = as.factor(cost), y = avg_dps, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Average DPS Comparison", x = "Cost", y = "Average DPS") +
  theme_minimal() +
  theme(legend.title = element_blank())

damage_comparison <- ggplot(combined_sets, aes(x = as.factor(cost), y = avg_damage, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Average Damage Comparison", x = "Cost", y = "Average Damage") +
  theme_minimal() +
  theme(legend.title = element_blank())

# Display interactive versions separately
plotly::ggplotly(health_comparison)
plotly::ggplotly(dps_comparison)
plotly::ggplotly(damage_comparison)
```

Bar charts indicate that higher costs correspond to higher stats across all star levels. However, the stats of 6-cost champions are significantly higher than those of any other cost category. Additionally, the stats for 1- to 5-cost champions in Set 12 tend to be slightly higher than those in Set 13, which makes the high stats of 6-cost champions even more pronounced. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Normalized stats Calculation
#############################################

# Calculate adjusted costs based on star level
arranged_champions_set12$adjusted_cost <- case_when(
  arranged_champions_set12$star_level == 1 ~ arranged_champions_set12$cost,
  arranged_champions_set12$star_level == 2 ~ arranged_champions_set12$cost * 3,
  arranged_champions_set12$star_level == 3 ~ arranged_champions_set12$cost * 9
)

arranged_champions_set13$adjusted_cost <- case_when(
  arranged_champions_set13$star_level == 1 ~ arranged_champions_set13$cost,
  arranged_champions_set13$star_level == 2 ~ arranged_champions_set13$cost * 3,
  arranged_champions_set13$star_level == 3 ~ arranged_champions_set13$cost * 9
)

# Calculate normalized values for Set 12
normalized_set12 <- arranged_champions_set12 %>%
  mutate(
    normalized_health = health / adjusted_cost,
    normalized_damage = damage / adjusted_cost,
    normalized_dps = dps / adjusted_cost
  )

# Calculate normalized values for Set 13
normalized_set13 <- arranged_champions_set13 %>%
  mutate(
    normalized_health = health / adjusted_cost,
    normalized_damage = damage / adjusted_cost,
    normalized_dps = dps / adjusted_cost
  )

# Average normalized values for Set 12
averaged_normalized_set12 <- normalized_set12 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_normalized_health = mean(normalized_health),
    avg_normalized_damage = mean(normalized_damage),
    avg_normalized_dps = mean(normalized_dps)
  )

# Average normalized values for Set 13
averaged_normalized_set13 <- normalized_set13 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_normalized_health = mean(normalized_health),
    avg_normalized_damage = mean(normalized_damage),
    avg_normalized_dps = mean(normalized_dps)
  )

# Add set labels
set12_normalized_avg_summary <- averaged_normalized_set12 %>%
  mutate(set = "Set 12")

set13_normalized_avg_summary <- averaged_normalized_set13 %>%
  mutate(set = "Set 13")

# Combine both sets into one dataset for comparison
combined_normalized_sets <- bind_rows(set12_normalized_avg_summary, set13_normalized_avg_summary)
```

## Set 12 vs Set 13: Normalized stats Comparison by Cost and Star Level

For these charts, stats are normalized by cost, where 1-star units are worth their base cost (1–6), 2-star units are worth three times their base cost, and 3-star units are worth nine times their base cost (Normalized Stat = Stat / Total Cost at Star Level). 

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=8}
#############################################
# Interactive Normalized Stats Comparison Visualizations
#############################################

# Comparison of normalized Health
norm_health_comparison <- ggplot(combined_normalized_sets, aes(x = as.factor(cost), y = avg_normalized_health, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Normalized Health Comparison", x = "Cost", y = "Normalized Health") +
  theme_minimal() +
  theme(legend.title = element_blank())

# Comparison of normalized DPS
norm_dps_comparison <- ggplot(combined_normalized_sets, aes(x = as.factor(cost), y = avg_normalized_dps, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Normalized DPS Comparison", x = "Cost", y = "Normalized DPS") +
  theme_minimal() +
  theme(legend.title = element_blank())

# Comparison of normalized Damage
norm_damage_comparison <- ggplot(combined_normalized_sets, aes(x = as.factor(cost), y = avg_normalized_damage, fill = set)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~star_level, nrow = 1, labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Normalized Damage Comparison", x = "Cost", y = "Normalized Damage") +
  theme_minimal() +
  theme(legend.title = element_blank())

# Display each as an interactive plot
plotly::ggplotly(norm_health_comparison)
plotly::ggplotly(norm_dps_comparison)
plotly::ggplotly(norm_damage_comparison)
```

The normalized bar charts show that although 6-cost champions have very high absolute stats, 1-cost champions offer a higher stats-to-cost ratio. When cost is taken into account, the normalized stats trend downward as cost increases, but the exceptionally high stats of 6-cost champions remain evident.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Heatmap Visualization for Normalized Health
#############################################

# Create heatmap for normalized health across sets and costs
heatmap_health <- ggplot(combined_normalized_sets, aes(x = as.factor(cost), y = as.factor(star_level), fill = avg_normalized_health)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "inferno") +
  labs(title = "Heatmap of Normalized Health", x = "Cost", y = "Star Level", fill = "Avg Norm Health") +
  theme_minimal() +
  facet_wrap(~set, nrow = 1)

print(heatmap_health)
```

The heatmap agrees with the bar charts above, supporting the claim that 1 costs have significantly higher normalized stats, namely health as shown by the heatmap, as compared to any other cost. 

## Set 12 Vs Set 13: Distribution Comparison

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Distribution Comparison with Violin Plots
#############################################

# Combine the datasets with an added 'set' column
combined_arranged <- bind_rows(
  arranged_champions_set12 %>% mutate(set = "Set 12"),
  arranged_champions_set13 %>% mutate(set = "Set 13")
)

# Create violin plots for health distribution
health_violin <- ggplot(combined_arranged, aes(x = as.factor(cost), y = health, fill = as.factor(star_level))) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  facet_grid(set ~ star_level, 
             labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Distribution of Health by Cost and Star Level Across Sets",
       x = "Cost",
       y = "Health") +
  theme_minimal() +
  theme(legend.position = "none")

# Create violin plots for DPS distribution
dps_violin <- ggplot(combined_arranged, aes(x = as.factor(cost), y = dps, fill = as.factor(star_level))) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  facet_grid(set ~ star_level, 
             labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Distribution of DPS by Cost and Star Level Across Sets",
       x = "Cost",
       y = "DPS") +
  theme_minimal() +
  theme(legend.position = "none")

# Create violin plots for damage distribution
damage_violin <- ggplot(combined_arranged, aes(x = as.factor(cost), y = damage, fill = as.factor(star_level))) +
  geom_violin(trim = FALSE, alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  facet_grid(set ~ star_level, 
             labeller = labeller(star_level = function(x) paste("★", x))) +
  labs(title = "Distribution of Damage by Cost and Star Level Across Sets",
       x = "Cost",
       y = "Damage") +
  theme_minimal() +
  theme(legend.position = "none")

# Print the plots
print(health_violin)
print(dps_violin)
print(damage_violin)
```

Violin plots reinforce previous trends, showing that higher-cost units with higher star levels generally have higher stats, though there are exceptions. Notably, the DPS of 3-star, 6-cost champions has a wide range, indicating that while some 6-cost champions exhibit very high DPS, others have much lower DPS. This variability highlights the unique impact of 6-cost champions in the game.