---
title: "Trait Win Rate and Model Summary"
output: 
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Install necessary packages (if not already installed)
# install.packages(c("rvest", "tidyverse", "stringr", "purrr"))
# install.packages("GGally")
# install.packages("plotly")
# install.packages("kableExtra")
# install.packages("randomForest")
# install.packages("DT")

library(plotly)
library(GGally)
library(rvest)
library(tidyverse)
library(dplyr)
library(purrr)
library(stringr)
library(jsonlite)
library(kableExtra)
library(tidyr)
library(ggplot2)
library(knitr)
library(gridExtra)
library(jsonlite)
library(tibble)
library(mgcv)
library(randomForest)
library(DT)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Wayback Machine Archive Data
#############################################

# Define the TFTTactics.gg Champions URL for Set 12
url_tfttactics_set12 <- "https://web.archive.org/web/20240927225144/https://tftactics.gg/champions/"
page_tfttactics_set12 <- read_html(url_tfttactics_set12)

# Extract champion names for Set 12
champion_names_set12 <- page_tfttactics_set12 %>%
  html_nodes(".character-name") %>% 
  html_text(trim = TRUE)

# Extract trait names (global list) for Set 12
trait_names_set12 <- page_tfttactics_set12 %>%
  html_nodes(".dropdown-item img") %>%
  html_attr("alt") 

# Initialize the tibbles with proper column names for Set 12
champion_info_set12 <- tibble(
  champion = character(),
  stats = list(),
  abilities = list(),
  traits = list(),
  trait_breakpoints = list()
)

trait_info_set12 <- tibble(
  champion = character(),
  trait_name = character(),
  trait_type = character(),
  trait_description = character()
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Wayback Machine Archive Data
#############################################

# Define the TFTTactics.gg Champions URL for Set 13
url_tfttactics_set13 <- "https://web.archive.org/web/20250114220710/https://tftactics.gg/champions/"
page_tfttactics_set13 <- read_html(url_tfttactics_set13)

# Extract champion names for Set 13
champion_names_set13 <- page_tfttactics_set13 %>%
  html_nodes(".character-name") %>% 
  html_text(trim = TRUE)

# Extract trait names (global list) for Set 13
trait_names_set13 <- page_tfttactics_set13 %>%
  html_nodes(".dropdown-item img") %>%
  html_attr("alt") 

# Initialize the tibbles with proper column names for Set 13
champion_info_set13 <- tibble(
  champion = character(),
  stats = list(),
  abilities = list(),
  traits = list(),
  trait_breakpoints = list()
)

trait_info_set13 <- tibble(
  champion = character(),
  trait_name = character(),
  trait_type = character(),
  trait_description = character()
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Common Functions (Used for both sets)
#############################################

# Function to safely read the page with error handling
safe_read_html <- function(url) {
  tryCatch(
    {
      read_html(url)
    },
    error = function(e) {
      message("Error reading URL: ", url)
      message("Error: ", e$message)
      NULL
    }
  )
}

# Function to extract champion data using regex patterns
extract_champion_data <- function(page, trait_names) {
  # Get raw HTML text
  raw_text <- page %>% html_text()
  
  # Extract champion name from the title
  champion_name <- page %>% 
    html_node("title") %>% 
    html_text() %>% 
    str_extract("^[^·]+") %>% 
    str_trim()
  
  # stats EXTRACTION
  stats <- tibble(stat = character(), value = character())
  stat_patterns <- list(
    "Cost" = "Cost:\\s*(\\d+)",
    "Health" = "Health:\\s*([\\d\\s\\/]+)",
    "Mana" = "Mana:\\s*(\\d+)",
    "Armor" = "Armor:\\s*(\\d+)",
    "MR" = "MR:\\s*(\\d+)",
    "DPS" = "DPS:\\s*([\\d\\s\\/]+)",
    "Damage" = "Damage:\\s*([\\d\\s\\/]+)",
    "Atk Spd" = "Atk Spd:\\s*([\\d\\.]+)",
    "Crit Rate" = "Crit Rate:\\s*([\\d\\.%]+)",
    "Range" = "Range:\\s*(\\d+)"
  )
  
  for (stat_name in names(stat_patterns)) {
    value <- str_extract(raw_text, stat_patterns[[stat_name]])
    if (!is.na(value)) {
      clean_value <- str_replace(value, paste0(stat_name, ":\\s*"), "") %>% str_trim()
      stats <- stats %>% add_row(stat = stat_name, value = clean_value)
    }
  }
  
  # ABILITIES EXTRACTION
  ability_section <- str_extract(raw_text, "Abilities(.+?)(?=[A-Z][a-z]+(?:Origin|Class)|Synergies|Privacy Policy)")
  ability_name <- str_extract(ability_section, "(?<=Abilities)\\s*([A-Za-z\\s]+?)(?=Active|Passive)") %>% str_trim()
  ability_type <- str_extract(ability_section, "(Active|Passive)")
  ability_mana <- str_extract(ability_section, "\\d+\\s*\\/\\s*\\d+")

  abilities <- tibble(
    name = ability_name,
    type = ability_type,
    mana = ability_mana,
    description = NA_character_
  )

  # TRAITS EXTRACTION
  champion_traits <- character()
  for (trait in trait_names) {
    if (str_detect(raw_text, fixed(trait))) {
      champion_traits <- c(champion_traits, trait)
    }
  }

  traits <- tibble(name = character(), type = character(), description = character())
  trait_breakpoints <- tibble(trait_name = character(), breakpoint_value = character(), breakpoint_description = character())

  for (trait_name in champion_traits) {
    next_traits <- setdiff(trait_names, trait_name)
    next_part <- if (length(next_traits) > 0) paste(next_traits, collapse = "|") else ""

    end_pattern <- if (nchar(next_part) > 0) {
      paste0("(?=", next_part, "|Synergies)")
    } else {
      "(?=Synergies)"
    }

    trait_pattern <- paste0(trait_name, "(.+?)", end_pattern)
    trait_section <- str_extract(raw_text, trait_pattern)

    if (!is.na(trait_section)) {
      trait_type <- str_extract(trait_section, "(Origin|Class)")

      desc_pattern <- paste0("(?<=", trait_type, ")\\s*(.+?)(?=(", 
                             if(nchar(next_part) > 0) paste0(next_part, "|") else "", 
                             "Synergies|\\d+%|\\d+\\s*AD))")

      trait_desc <- str_extract(trait_section, desc_pattern) %>% str_trim()

      traits <- traits %>% add_row(
        name = trait_name,
        type = trait_type,
        description = trait_desc
      )

      breakpoint_pattern <- "(\\d+)(.+?)(?=\\d+|$)"
      breakpoint_matches <- str_extract_all(trait_section, breakpoint_pattern)[[1]]
      for (match in breakpoint_matches) {
        bp_value <- str_extract(match, "^\\d+") %>% str_trim()
        bp_desc <- str_replace(match, "^\\d+", "") %>% str_trim()
        trait_breakpoints <- trait_breakpoints %>% add_row(
          trait_name = trait_name,
          breakpoint_value = bp_value,
          breakpoint_description = bp_desc
        )
      }
    }
  }

  return(list(
    champion = champion_name,
    stats = stats,
    abilities = abilities,
    traits = traits,
    trait_breakpoints = trait_breakpoints
  ))
}

# Function to extract champion stats for analysis
extract_champion_stats <- function(champion_name, stats_tibble) {
  # Extract Cost, Health, DPS, and Damage stats
  cost <- stats_tibble %>% filter(stat == "Cost") %>% pull(value) %>% as.numeric()
  health <- stats_tibble %>% filter(stat == "Health") %>% pull(value)
  dps <- stats_tibble %>% filter(stat == "DPS") %>% pull(value)
  damage <- stats_tibble %>% filter(stat == "Damage") %>% pull(value)
  
  # Split the health, dps, and damage values by " / " to separate star levels
  health_values <- as.numeric(strsplit(health, " / ")[[1]])
  dps_values <- as.numeric(strsplit(dps, " / ")[[1]])
  damage_values <- as.numeric(strsplit(damage, " / ")[[1]])

  # Ensure that the number of values for health, dps, and damage match across star levels
  if (length(health_values) != length(dps_values) | length(dps_values) != length(damage_values)) {
    stop("Mismatch in the number of star levels for health, dps, and damage.")
  }
  
  # Create a tibble with the expanded stats
  stats_df <- tibble(
    champion = rep(champion_name, length(health_values)),
    star_level = 1:length(health_values),
    cost = rep(cost, length(health_values)),
    health = health_values,
    dps = dps_values,
    damage = damage_values
  )
  
  return(stats_df)
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Extract All Champion Data
#############################################

extract_all_champion_data_set12 <- function(champion_names) {
  # Reset global variables
  champion_info_set12 <<- tibble(
    champion = character(),
    stats = list(),
    abilities = list(),
    traits = list(),
    trait_breakpoints = list()
  )

  trait_info_set12 <<- tibble(
    champion = character(),
    trait_name = character(),
    trait_type = character(),
    trait_description = character()
  )

  base_url <- "https://web.archive.org/web/20240927225144/https://tftactics.gg/champions/"

  for (champion_name in champion_names) {
    champion_url <- paste0(base_url, gsub(" ", "_", champion_name), "/")
    champion_page <- safe_read_html(champion_url)

    if (is.null(champion_page)) {
      next
    }

    champion_data <- extract_champion_data(champion_page, trait_names_set12)

    champion_info_set12 <<- champion_info_set12 %>% add_row(
      champion = champion_data$champion,
      stats = list(champion_data$stats),
      abilities = list(champion_data$abilities),
      traits = list(champion_data$traits),
      trait_breakpoints = list(champion_data$trait_breakpoints)
    )

    for (trait in champion_data$traits$name) {
      trait_info_set12 <<- trait_info_set12 %>% add_row(
        champion = champion_data$champion,
        trait_name = trait,
        trait_type = champion_data$traits$type[champion_data$traits$name == trait],
        trait_description = champion_data$traits$description[champion_data$traits$name == trait]
      )
    }
  }

  return(list(champion_info = champion_info_set12, trait_info = trait_info_set12))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Extract All Champion Data
#############################################

extract_all_champion_data_set13 <- function(champion_names) {
  # Reset global variables
  champion_info_set13 <<- tibble(
    champion = character(),
    stats = list(),
    abilities = list(),
    traits = list(),
    trait_breakpoints = list()
  )

  trait_info_set13 <<- tibble(
    champion = character(),
    trait_name = character(),
    trait_type = character(),
    trait_description = character()
  )

  base_url <- "https://tftactics.gg/champions/"

  for (champion_name in champion_names) {
    champion_url <- paste0(base_url, gsub(" ", "_", champion_name), "/")
    champion_page <- safe_read_html(champion_url)

    if (is.null(champion_page)) {
      next
    }

    champion_data <- extract_champion_data(champion_page, trait_names_set13)

    champion_info_set13 <<- champion_info_set13 %>% add_row(
      champion = champion_data$champion,
      stats = list(champion_data$stats),
      abilities = list(champion_data$abilities),
      traits = list(champion_data$traits),
      trait_breakpoints = list(champion_data$trait_breakpoints)
    )

    for (trait in champion_data$traits$name) {
      trait_info_set13 <<- trait_info_set13 %>% add_row(
        champion = champion_data$champion,
        trait_name = trait,
        trait_type = champion_data$traits$type[champion_data$traits$name == trait],
        trait_description = champion_data$traits$description[champion_data$traits$name == trait]
      )
    }
  }

  return(list(champion_info = champion_info_set13, trait_info = trait_info_set13))
}
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Extract Data for Both Sets
#############################################
# Extract data for both sets
# all_data_set12 <- extract_all_champion_data_set12(champion_names_set12)
# all_data_set13 <- extract_all_champion_data_set13(champion_names_set13)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read the JSON files
champion_info_set12 <- fromJSON("data/champion_info_set12.json")
trait_info_set12 <- fromJSON("data/trait_info_set12.json")

# Initialize all_data_set_12 as an empty data frame
all_data_set12 <- list()

# Add the JSON data as columns to the new data frame
all_data_set12$champion_info <- champion_info_set12
all_data_set12$trait_info <- trait_info_set12
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Read the JSON files
champion_info_set13 <- fromJSON("data/champion_info_set13.json")
trait_info_set13 <- fromJSON("data/trait_info_set13.json")

# Initialize all_data_set_13 as an empty data frame
all_data_set13 <- list()

# Add the JSON data as columns to the new data frame
all_data_set13$champion_info <- champion_info_set13
all_data_set13$trait_info <- trait_info_set13
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Process Data for Both Sets
#############################################
# Clean champion names for Set 12
all_data_set12$trait_info <- all_data_set12$trait_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide")) %>%
  filter(!(trait_name == "Portal" & is.na(trait_type)))

all_data_set12$champion_info <- all_data_set12$champion_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

# Clean champion names for Set 13
all_data_set13$trait_info <- all_data_set13$trait_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

all_data_set13$champion_info <- all_data_set13$champion_info %>%
  mutate(champion = str_remove_all(champion, "TFT | Guide"))

# print(all_data_set12$trait_info)
# print(all_data_set12$champion_info)
# print(all_data_set13$trait_info)
# print(all_data_set13$champion_info)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 12 - Extract All Winrate Data
#############################################

# Read JSON
url_set12 <- "https://web.archive.org/web/20240917151936/https://tftactics-data.kda.gg/meta/report"
report_data_set12 <- fromJSON(url_set12)

# Go to the summary-section
summary_section_set12 <- report_data_set12$data$`summary-section`

# Helper function to clean a single name
clean_champion_name_set12 <- function(x) {
  x <- gsub("^[0-9]-", "", x)
  x <- gsub("^tft12_", "", x)
  x <- gsub("-[0-9]+$", "", x)
  x <- tolower(x)
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x
}

# Build the dataframe
comp_df_set12 <- tibble(
  win_rate = summary_section_set12$win_rate,
  top4_rate = summary_section_set12$top4_rate,
  champions = sapply(summary_section_set12$composition, function(champs_vec) {
    cleaned_champions <- sapply(champs_vec, clean_champion_name_set12)
    paste(cleaned_champions, collapse = ", ")
  })
)

# Separate Champions from Traits
comp_df_clean_set12 <- comp_df_set12 %>%
  mutate(champions = str_split(champions, ",\\s*")) %>%
  rowwise() %>%
  mutate(
    champions_list = list(champions[champions %in% champion_names_set12]),
    traits_list = list(champions[!(champions %in% champion_names_set12)])
  ) %>%
  ungroup() %>%
  mutate(
    champions = sapply(champions_list, function(x) paste(x, collapse = ", ")),
    traits = sapply(traits_list, function(x) paste(x, collapse = ", "))
  ) %>%
  select(win_rate, top4_rate, champions, traits)

# print(comp_df_clean_set12)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# SET 13 - Extract All Winrate Data
#############################################

# Download compressed file
temp_file_set13 <- tempfile(fileext = ".gz")
download.file(
  url = "https://web.archive.org/web/20250113142238/https://tftactics-data.kda.gg/meta/report",
  destfile = temp_file_set13,
  mode = "wb"
)

# Decompress using gzcon
con_set13 <- gzcon(file(temp_file_set13, "rb"))
text_set13 <- readLines(con_set13, warn = FALSE)
close(con_set13)

# Parse the text as JSON
report_data_set13 <- fromJSON(paste(text_set13, collapse = ""))

# Go to the summary-section
summary_section_set13 <- report_data_set13$data$`summary-section`

# Helper function to clean a single name
clean_champion_name_set13 <- function(x) {
  x <- gsub("^[0-9]-", "", x)
  x <- gsub("^tft13_", "", x)
  x <- gsub("-[0-9]+$", "", x)
  x <- tolower(x)
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x
}

# Build the dataframe
comp_df_set13 <- tibble(
  win_rate = summary_section_set13$win_rate,
  top4_rate = summary_section_set13$top4_rate,
  champions = sapply(summary_section_set13$composition, function(champs_vec) {
    cleaned_champions <- sapply(champs_vec, clean_champion_name_set13)
    paste(cleaned_champions, collapse = ", ")
  })
)

# Separate Champions from Traits
comp_df_clean_set13 <- comp_df_set13 %>%
  mutate(champions = str_split(champions, ",\\s*")) %>%
  rowwise() %>%
  mutate(
    champions_list = list(champions[champions %in% champion_names_set13]),
    traits_list = list(champions[!(champions %in% champion_names_set13)])
  ) %>%
  ungroup() %>%
  mutate(
    champions = sapply(champions_list, function(x) paste(x, collapse = ", ")),
    traits = sapply(traits_list, function(x) paste(x, collapse = ", "))
  ) %>%
  select(win_rate, top4_rate, champions, traits)

# print(comp_df_clean_set13)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 12 and Set 13 Champion and Trait Stats
#############################################

# Explode champions and traits into separate rows

# For Set 12 - explode champions
comp_df_clean_set12_exploded <- comp_df_clean_set12 %>%
  mutate(
    champions = str_split(champions, ",\\s*")
  ) %>%
  unnest(champions) %>%
  mutate(
    champions = str_trim(champions)
  )

# For Set 13 - explode champions
comp_df_clean_set13_exploded <- comp_df_clean_set13 %>%
  mutate(
    champions = str_split(champions, ",\\s*")
  ) %>%
  unnest(champions) %>%
  mutate(
    champions = str_trim(champions)
  )

# For Set 12 - explode traits
comp_df_clean_set12_traits_exploded <- comp_df_clean_set12 %>%
  mutate(
    traits = str_split(traits, ",\\s*")
  ) %>%
  unnest(traits) %>%
  mutate(
    traits = str_trim(traits)
  )

# For Set 13 - explode traits
comp_df_clean_set13_traits_exploded <- comp_df_clean_set13 %>%
  mutate(
    traits = str_split(traits, ",\\s*")
  ) %>%
  unnest(traits) %>%
  mutate(
    traits = str_trim(traits)
  )

# Summarize Champion Stats
champion_summary_set12 <- comp_df_clean_set12_exploded %>%
  group_by(champions) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

champion_summary_set13 <- comp_df_clean_set13_exploded %>%
  group_by(champions) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# Summarize Trait Stats
trait_summary_set12 <- comp_df_clean_set12_traits_exploded %>%
  group_by(traits) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

trait_summary_set13 <- comp_df_clean_set13_traits_exploded %>%
  group_by(traits) %>%
  summarise(
    avg_win_rate = mean(win_rate, na.rm = TRUE),
    avg_top4_rate = mean(top4_rate, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# champion_summary_set12
# champion_summary_set13
# trait_summary_set12
# trait_summary_set13
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 12 Champion stats
#############################################

# Prepare champion stats for analysis (Set 12)
all_champion_stats_set12 <- bind_rows(
  lapply(1:nrow(all_data_set12$champion_info), function(i) {
    champion_name <- all_data_set12$champion_info$champion[i]
    champion_stats <- all_data_set12$champion_info$stats[[i]]
    
    # Extract and combine stats
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
      # return(NULL)
    })
  })
)

# Remove any NULL entries
all_champion_stats_set12 <- all_champion_stats_set12 %>% filter(!is.na(champion))

# Arrange champions by various attributes (Set 12)
arranged_champions_set12 <- all_champion_stats_set12 %>%
  arrange(champion, star_level, cost)

# Compute averages for health, dps, and damage grouped by cost and star level (Set 12)
averaged_champions_set12 <- arranged_champions_set12 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping after summarisation
  )

# Sort tables by different metrics (Set 12)
ordered_by_health_set12 <- averaged_champions_set12 %>%
  arrange(avg_health) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_dps_set12 <- averaged_champions_set12 %>%
  arrange(avg_dps) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_damage_set12 <- averaged_champions_set12 %>%
  arrange(avg_damage) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

# Create kable tables for Set 12
kable_health_set12 <- kable(ordered_by_health_set12, 
                          caption = "Set 12: Champions Ordered by Average Health",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

kable_dps_set12 <- kable(ordered_by_dps_set12, 
                       caption = "Set 12: Champions Ordered by Average DPS",
                       col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                       format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

kable_damage_set12 <- kable(ordered_by_damage_set12, 
                          caption = "Set 12: Champions Ordered by Average Damage",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#3498DB")

# print(kable_health_set12)
# print(kable_dps_set12)
# print(kable_damage_set12)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Analyze Set 13 Champion stats
#############################################

# Prepare champion stats for analysis (Set 13)
all_champion_stats_set13 <- bind_rows(
  lapply(1:nrow(all_data_set13$champion_info), function(i) {
    champion_name <- all_data_set13$champion_info$champion[i]
    champion_stats <- all_data_set13$champion_info$stats[[i]]
    
    # Extract and combine stats
    tryCatch({
      extract_champion_stats(champion_name, champion_stats)
    }, error = function(e) {
      message(paste("Error processing champion:", champion_name))
      message(e$message)
      # return(NULL)
    })
  })
)

# Remove any NULL entries
all_champion_stats_set13 <- all_champion_stats_set13 %>% filter(!is.na(champion))

# Arrange champions by various attributes (Set 13)
arranged_champions_set13 <- all_champion_stats_set13 %>%
  arrange(champion, star_level, cost)

# Compute averages for health, dps, and damage grouped by cost and star level (Set 13)
averaged_champions_set13 <- arranged_champions_set13 %>%
  group_by(cost, star_level) %>%
  summarise(
    avg_health = mean(health, na.rm = TRUE),
    avg_dps = mean(dps, na.rm = TRUE),
    avg_damage = mean(damage, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping after summarisation
  )

# Sort tables by different metrics (Set 13)
ordered_by_health_set13 <- averaged_champions_set13 %>%
  arrange(avg_health) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_dps_set13 <- averaged_champions_set13 %>%
  arrange(avg_dps) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

ordered_by_damage_set13 <- averaged_champions_set13 %>%
  arrange(avg_damage) %>%
  mutate(
    avg_health = round(avg_health, 1),
    avg_dps = round(avg_dps, 1),
    avg_damage = round(avg_damage, 1)
  )

# Create kable tables for Set 13
kable_health_set13 <- kable(ordered_by_health_set13, 
                          caption = "Set 13: Champions Ordered by Average Health",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

kable_dps_set13 <- kable(ordered_by_dps_set13, 
                       caption = "Set 13: Champions Ordered by Average DPS",
                       col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                       format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

kable_damage_set13 <- kable(ordered_by_damage_set13, 
                          caption = "Set 13: Champions Ordered by Average Damage",
                          col.names = c("Cost", "Star Level", "Avg Health", "Avg DPS", "Avg Damage"),
                          format = "html") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#E74C3C")

# print(kable_health_set13)
# print(kable_dps_set13)
# print(kable_damage_set13)
```

# Champion and Trait Win Rate Analysis

So far, we've explored how champion stats like health, damage, and DPS scale across costs and star levels in Sets 12 and 13. However, high stats don't necessarily translate to better performance. To assess how effective champions and traits actually are in-game, we now incorporate **win rate** as our performance metric.

Win rate in this analysis refers to the **Top 1 Rate**, derived from thousands of real team compositions. Each champion's win rate is averaged across all compositions in which it appears. This allows us to examine whether certain champions or traits consistently contribute to victory — regardless of raw stats.

The following interactive visualizations explore:
- How champion cost relates to average win rate  
- Which traits are most strongly associated with winning teams

## Champion Win Rate vs Cost

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Scatterplot: Cost vs Win Rate
#############################################

# Combine champion summaries
scatter_data <- bind_rows(
  champion_summary_set12 %>%
    mutate(set = "Set 12"),
  champion_summary_set13 %>%
    mutate(set = "Set 13")
)

# Add cost info from stat data (assuming 3-star only)
cost_lookup <- bind_rows(arranged_champions_set12, arranged_champions_set13) %>%
  filter(star_level == 3) %>%
  select(champion, cost) %>%
  distinct()

# Merge cost and scatter_data
scatter_data <- scatter_data %>%
  left_join(cost_lookup, by = c("champions" = "champion"))

# Interactive Plot
plot_ly(
  data = scatter_data,
  x = ~cost,
  y = ~avg_win_rate,
  type = "scatter",
  mode = "markers",
  color = ~set,
  text = ~paste("Champion:", champions,
                "<br>Set:", set,
                "<br>Cost:", cost,
                "<br>Avg Win Rate:", round(avg_win_rate, 4)),
  hoverinfo = "text"
) %>%
  layout(
    title = "Champion Win Rate vs Cost",
    xaxis = list(title = "Champion Cost"),
    yaxis = list(title = "Average Win Rate"),
    legend = list(title = list(text = "<b>Set</b>"))
  )
```

We can see from the scatterplot that the winrates of the best performing champions in Set 12 are higher than that of Set 13. This could indicate that the balancing is better in Set 13. 

## Trait-Level Win Rate Summary

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Trait Win Rate Table (Set 12 & 13)
#############################################

# Combine both trait name lists
valid_trait_names <- union(trait_names_set12, trait_names_set13)

# Combine and filter both trait summaries
trait_table_data <- bind_rows(
  trait_summary_set12 %>%
    mutate(set = "Set 12"),
  trait_summary_set13 %>%
    mutate(set = "Set 13")
) %>%
  filter(traits %in% valid_trait_names) %>%
  mutate(
    avg_win_rate = round(avg_win_rate, 4),
    avg_top4_rate = round(avg_top4_rate, 4)
  ) %>%
  arrange(desc(avg_win_rate))

# Display the interactive table
datatable(
  trait_table_data,
  options = list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE,
  colnames = c("Trait", "Avg Win Rate", "Avg Top 4 Rate", "Count", "Set"),
  caption = "Interactive Table: Trait-Level Win Rate Summary"
)
```

## Predicting Win Rate: Stats, Traits, and Models

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Regression: Champion Stats vs. Win Rate & Top 4 Rate
#############################################

# Average stats across star levels per champion for Set 12
champion_avg_stats_set12 <- arranged_champions_set12 %>%
  group_by(champion) %>%
  summarise(
    cost = mean(cost, na.rm = TRUE),
    health = mean(health, na.rm = TRUE),
    damage = mean(damage, na.rm = TRUE),
    dps = mean(dps, na.rm = TRUE),
    .groups = "drop"
  )

# Average stats across star levels per champion for Set 13
champion_avg_stats_set13 <- arranged_champions_set13 %>%
  group_by(champion) %>%
  summarise(
    cost = mean(cost, na.rm = TRUE),
    health = mean(health, na.rm = TRUE),
    damage = mean(damage, na.rm = TRUE),
    dps = mean(dps, na.rm = TRUE),
    .groups = "drop"
  )

# Join averaged stats with composition data (Set 12)
regression_df_set12 <- comp_df_clean_set12_exploded %>%
  left_join(champion_avg_stats_set12, by = c("champions" = "champion")) %>%
  filter(!is.na(health), !is.na(dps), !is.na(damage)) %>%
  mutate(set = "Set 12")

# Join averaged stats with composition data (Set 13)
regression_df_set13 <- comp_df_clean_set13_exploded %>%
  left_join(champion_avg_stats_set13, by = c("champions" = "champion")) %>%
  filter(!is.na(health), !is.na(dps), !is.na(damage)) %>%
  mutate(set = "Set 13")
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Prepare Trait Dummies and Merge with Stats (Set 12 & Set 13)
#############################################

# Clean and combine traits into dummy variables per champion (Set 12)
trait_dummies_set12 <- all_data_set12$trait_info %>%
  mutate(trait_name = make.names(trait_name)) %>%  # Handles spaces, punctuation
  select(champion, trait_name) %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = trait_name, values_from = value, values_fill = 0)

# Clean and combine traits into dummy variables per champion (Set 13)
trait_dummies_set13 <- all_data_set13$trait_info %>%
  mutate(trait_name = make.names(trait_name)) %>%
  select(champion, trait_name) %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = trait_name, values_from = value, values_fill = 0)

# Merge stats + traits for Set 12
regression_full_set12 <- regression_df_set12 %>%
  left_join(trait_dummies_set12, by = c("champions" = "champion"))

# Merge stats + traits for Set 13
regression_full_set13 <- regression_df_set13 %>%
  left_join(trait_dummies_set13, by = c("champions" = "champion"))

# Helper function: remove columns with <2 unique values (no variation)
drop_constant_columns <- function(df) {
  df %>%
    select(where(~ n_distinct(.) > 1)) %>%
    drop_na()
}

# Traits only (drop stats) and drop constant traits
regression_traits_only_set12 <- regression_full_set12 %>%
  select(-c(cost, health, damage, dps)) %>%
  drop_constant_columns()

regression_traits_only_set13 <- regression_full_set13 %>%
  select(-c(cost, health, damage, dps)) %>%
  drop_constant_columns()

# Full set (stats + traits) with constant traits dropped
regression_full_set12 <- regression_full_set12 %>%
  drop_constant_columns()

regression_full_set13 <- regression_full_set13 %>%
  drop_constant_columns()
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Linear Models (LM) - Win Rate ~ Champion Stats, Traits, and Both
#############################################

# Set 12 - Linear Models
lm_stats_set12 <- lm(win_rate ~ cost + health + damage + dps, data = regression_df_set12)
lm_traits_set12 <- lm(win_rate ~ ., data = regression_traits_only_set12 %>% select(-top4_rate))
lm_both_set12 <- lm(win_rate ~ ., data = regression_full_set12 %>% select(-top4_rate))

# Set 13 - Linear Models
lm_stats_set13 <- lm(win_rate ~ cost + health + damage + dps, data = regression_df_set13)
lm_traits_set13 <- lm(win_rate ~ ., data = regression_traits_only_set13 %>% select(-top4_rate))
lm_both_set13 <- lm(win_rate ~ ., data = regression_full_set13 %>% select(-top4_rate))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Helper: Safe Formula Constructor (With Backticks)
#############################################

safe_formula_string <- function(df, response, exclude = NULL) {
  predictors <- df %>%
    select(where(is.numeric)) %>%               # Keep only numeric columns
    select(-all_of(c(response, exclude))) %>%   # Drop response and excluded
    names()                                     # Get column names
  
  predictors_wrapped <- paste0("`", predictors, "`")  # Backtick everything
  as.formula(paste(response, "~", paste(predictors_wrapped, collapse = " + ")))
}
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Generalized Additive Models (GAM) - Win Rate ~ Champion Stats, Traits, and Both
#############################################

# Set 12
gam_stats_set12 <- gam(win_rate ~ s(cost, k = 4) + s(health) + s(damage) + s(dps), data = regression_df_set12)

gam_traits_set12 <- gam(
  formula = safe_formula_string(regression_traits_only_set12, "win_rate", exclude = "top4_rate"),
  data = regression_traits_only_set12
)

gam_both_set12 <- gam(
  formula = safe_formula_string(regression_full_set12, "win_rate", exclude = "top4_rate"),
  data = regression_full_set12
)

# Set 13
gam_stats_set13 <- gam(win_rate ~ s(cost, k = 4) + s(health) + s(damage) + s(dps), data = regression_df_set13)

gam_traits_set13 <- gam(
  formula = safe_formula_string(regression_traits_only_set13, "win_rate", exclude = "top4_rate"),
  data = regression_traits_only_set13
)

gam_both_set13 <- gam(
  formula = safe_formula_string(regression_full_set13, "win_rate", exclude = "top4_rate"),
  data = regression_full_set13
)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Random Forest (RF) - Win Rate ~ Champion Stats, Traits, and Both
#############################################

# Set 12 - RF
rf_stats_set12 <- randomForest(win_rate ~ cost + health + damage + dps, data = regression_df_set12, ntree = 500)
rf_traits_set12 <- randomForest(win_rate ~ ., data = regression_traits_only_set12 %>% select(-top4_rate), ntree = 500)
rf_both_set12 <- randomForest(win_rate ~ ., data = regression_full_set12 %>% select(-top4_rate), ntree = 500)

# Set 13 - RF
rf_stats_set13 <- randomForest(win_rate ~ cost + health + damage + dps, data = regression_df_set13, ntree = 500)
rf_traits_set13 <- randomForest(win_rate ~ ., data = regression_traits_only_set13 %>% select(-top4_rate), ntree = 500)
rf_both_set13 <- randomForest(win_rate ~ ., data = regression_full_set13 %>% select(-top4_rate), ntree = 500)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Model Comparison Table - R², RMSE, AIC
#############################################

model_comparison <- tibble::tibble(
  Model_Type = rep(c("LM", "GAM", "RF"), each = 6),
  Set = rep(rep(c("Set 12", "Set 13"), each = 3), times = 3),
  Input = rep(c("Stats only", "Traits only", "Both"), times = 6),

  R2 = c(
    # LM
    summary(lm_stats_set12)$adj.r.squared,
    summary(lm_traits_set12)$adj.r.squared,
    summary(lm_both_set12)$adj.r.squared,
    summary(lm_stats_set13)$adj.r.squared,
    summary(lm_traits_set13)$adj.r.squared,
    summary(lm_both_set13)$adj.r.squared,

    # GAM
    summary(gam_stats_set12)$r.sq,
    summary(gam_traits_set12)$r.sq,
    summary(gam_both_set12)$r.sq,
    summary(gam_stats_set13)$r.sq,
    summary(gam_traits_set13)$r.sq,
    summary(gam_both_set13)$r.sq,

    # RF
    tail(rf_stats_set12$rsq, 1),
    tail(rf_traits_set12$rsq, 1),
    tail(rf_both_set12$rsq, 1),
    tail(rf_stats_set13$rsq, 1),
    tail(rf_traits_set13$rsq, 1),
    tail(rf_both_set13$rsq, 1)
  ),

  RMSE = c(
    # LM
    sqrt(mean(residuals(lm_stats_set12)^2)),
    sqrt(mean(residuals(lm_traits_set12)^2)),
    sqrt(mean(residuals(lm_both_set12)^2)),
    sqrt(mean(residuals(lm_stats_set13)^2)),
    sqrt(mean(residuals(lm_traits_set13)^2)),
    sqrt(mean(residuals(lm_both_set13)^2)),

    # GAM
    sqrt(mean(residuals(gam_stats_set12)^2)),
    sqrt(mean(residuals(gam_traits_set12)^2)),
    sqrt(mean(residuals(gam_both_set12)^2)),
    sqrt(mean(residuals(gam_stats_set13)^2)),
    sqrt(mean(residuals(gam_traits_set13)^2)),
    sqrt(mean(residuals(gam_both_set13)^2)),

    # RF
    sqrt(mean(rf_stats_set12$mse)),
    sqrt(mean(rf_traits_set12$mse)),
    sqrt(mean(rf_both_set12$mse)),
    sqrt(mean(rf_stats_set13$mse)),
    sqrt(mean(rf_traits_set13$mse)),
    sqrt(mean(rf_both_set13$mse))
  ),

  AIC = c(
    # LM
    AIC(lm_stats_set12),
    AIC(lm_traits_set12),
    AIC(lm_both_set12),
    AIC(lm_stats_set13),
    AIC(lm_traits_set13),
    AIC(lm_both_set13),

    # GAM
    AIC(gam_stats_set12),
    AIC(gam_traits_set12),
    AIC(gam_both_set12),
    AIC(gam_stats_set13),
    AIC(gam_traits_set13),
    AIC(gam_both_set13),

    # RF — not defined
    rep(NA, 6)
  )
)

# Display
knitr::kable(model_comparison, digits = 4, format = "html", caption = "Model Comparison: R², RMSE, AIC (if applicable)") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

Random Forest with traits only is the best model overall because it has the highest R^2 and lowest RMSE across both sets, indicating strong predictive performance. Linear models likely overfit, as shown by perfect R^2 and zero RMSE, while GAMs perform worse in both R² and RMSE without gains in AIC. Traits only outperforms stats only and both combined, likely because stats add noise or redundant information.

## Trait Importance Analysis

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#############################################
# Random Forest (RF) - Trait Importance for Win Rate and Top 4 Rate
#############################################

# Helper function: Fit Random Forest and extract top 15 variable importances
get_top_traits_clean <- function(df, response_var, remove_var) {
  clean_df <- df %>%
    select(-traits, -champions, -all_of(remove_var)) %>%
    drop_na()

  rf_model <- randomForest(
    as.formula(paste(response_var, "~ .")),
    data = clean_df,
    ntree = 500
  )

  imp_vals <- importance(rf_model)
  imp_df <- data.frame(Trait = rownames(imp_vals), Importance = imp_vals[, 1])
  top15 <- imp_df %>%
    arrange(desc(Importance)) %>%
    head(15)
  return(top15)
}

# Extract top traits (removing the other target variable!)
top15_set12_winrate <- get_top_traits_clean(regression_traits_only_set12, "win_rate", "top4_rate")
top15_set12_top4rate <- get_top_traits_clean(regression_traits_only_set12, "top4_rate", "win_rate")
top15_set13_winrate <- get_top_traits_clean(regression_traits_only_set13, "win_rate", "top4_rate")
top15_set13_top4rate <- get_top_traits_clean(regression_traits_only_set13, "top4_rate", "win_rate")

# Helper function to generate a plot
make_plot <- function(df, set_label, metric_label, fill_color) {
  ggplot(df, aes(x = reorder(Trait, Importance), y = Importance)) +
    geom_col(fill = fill_color) +
    coord_flip() +
    labs(
      title = paste("Top 15 Traits for", metric_label, "-", set_label),
      x = "Trait",
      y = "Importance"
    ) +
    theme_minimal()
}

# Create each plot
plot1 <- make_plot(top15_set12_winrate, "Set 12", "Win Rate", "#2E86C1")
plot2 <- make_plot(top15_set12_top4rate, "Set 12", "Top 4 Rate", "#2E86C1")
plot3 <- make_plot(top15_set13_winrate, "Set 13", "Win Rate", "#C0392B")
plot4 <- make_plot(top15_set13_top4rate, "Set 13", "Top 4 Rate", "#C0392B")

# Display plots in 2x2 layout
grid.arrange(plot1, plot2, plot3, plot4, ncol = 2)
```

From the figure above, the best traits for Set 12 are Vanguard, Blaster and Honeymancy, while the best traits for Set 13 are Artillerist, Ambusher and Experiment. 

# Summary

This project investigated how champion stats, traits, and game design changes between Set 12 and Set 13 in *Teamfight Tactics* (TFT) influence team performance. By analyzing thousands of real ranked team compositions and applying multiple modeling approaches, we identified key structural changes and predictors of success in the evolving TFT metagame.

## Key Findings

- **Champion Stat Scaling**  
  Champion stats (Health, Damage, DPS) scale positively with both cost and star level in both sets. However, Set 13 introduces a **sharper stat gradient**, primarily driven by the addition of **6-cost champions**, which significantly outpace lower-cost units in raw stats. This supports the hypothesis that Set 13 meaningfully expanded the power curve.

- **Normalized Stat Efficiency**  
  After adjusting for the gold required to upgrade units (i.e., total cost at each star level), **1-cost champions provide the best stat-per-gold efficiency**. Normalized stats decline as unit cost increases, suggesting that **higher-cost champions are less efficient**, despite being more powerful in absolute terms.

- **Stat Progression Consistency**  
  Stat progression across star levels is remarkably consistent across sets, indicating that **design shifts in Set 13 built upon existing scaling rules** rather than redefining them. The key difference lies in the expanded ceiling rather than altered stat progression logic.

- **Trait Dominance over Raw Stats**  
  Predictive models consistently identified **traits as more important than raw stats** in explaining win rate and top-four rate. This underscores the central role of **trait synergies**, which amplify performance through interactions between units, rather than individual power alone.

- **Model Performance and Complexity**  
  Among all model types tested—Linear Models (LM), Generalized Additive Models (GAM), and Random Forests (RF)—**Random Forest models using trait-only features performed best**, achieving the highest $R^2$ and lowest RMSE. This confirms that **nonlinear, interaction-based models are better suited** for capturing the combinatorial complexity of TFT.

- **Top Traits by Set**  
  Trait importance varied across sets, reflecting shifting metas and design goals:  
  - *Set 12*: Vanguard, Blaster, and Honeymancy were key contributors to winning teams.  
  - *Set 13*: Artillerist, Ambusher, and Experiment emerged as the most influential traits.

## Overall Conclusion

The introduction of 6-cost champions in Set 13 marked a significant escalation in power, reshaping the competitive landscape. However, this power spike did not break existing game balance mechanisms; rather, it **layered new strategic depth onto the existing framework**. Across both sets, **traits remained the dominant factor in team success**, validating the game’s emphasis on synergy over raw strength.

## Broader Implications and Limitations

This analysis highlights the strength of trait-based design in sustaining strategic variety across TFT sets. However, it is limited by its reliance on **composition-level data**, which prevents precise measurement of individual unit contributions. Moreover, model predictions are constrained by the **accuracy and completeness of scraped data** from third-party sources. Future analyses could benefit from player-level match data or official Riot APIs to improve granularity and causal inference.
